---
title: "Full Replication: CEO--Employee Political Misalignment and Corporate Restructuring"
author: "Sina Davoudi"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
    toc: true
    toc_depth: 2
    number_sections: true
geometry: margin=0.8in
fontsize: 10pt
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \floatplacement{table}{H}
  - \usepackage{caption}
  - \captionsetup{font=small, labelfont=bf, skip=5pt}
  - \usepackage{threeparttable}
  - \usepackage{setspace}
  - \singlespacing
  - \pagestyle{plain}
editor_options: 
  markdown: 
    wrap: 72
---

\newpage

# Section 0: Packages and Setup

This chunk loads all R packages used throughout the replication and sets
global options.

```{r}
knitr::opts_chunk$set(echo = TRUE)
## ── Packages ──
library(data.table)
library(fixest)
library(kableExtra)
library(stringdist)

## Suppress fixest estimation notes in output
setFixest_notes(FALSE)

## Set working directory for all notebook chunks
knitr::opts_knit$set(
  root.dir = "/Users/homemac/Library/Mobile Documents/com~apple~CloudDocs/My Research/Accounting Research/Political_Accounting/Political Misalignment/Codes"
)
```

\newpage

# Section 1: Cycle Processing Function {#sec-cycle}

This section contains the reusable function that converts a raw DIME
ContribDB cycle file (e.g., `contribDB_2018.csv`) into a firm-level
employee ideology panel (`employee_panel_YYYY.rdata`).

**How it works (pre-filter approach for speed):**

1.  Read only the 5 needed columns from the CSV (keeps RAM low).
2.  Filter to individual contributors with a valid CFscore and employer.
3.  **Step 1 -- fast pre-match:** uppercase + trim employer names, then
    match directly against the employer lookup. This catches \~84 % of
    matchable rows *without* any regex.
4.  **Step 2 -- suffix cleaning on unmatched remainder only:** run the
    expensive `gsub()` suffix-removal loop on the \~16 % of unmatched
    rows, then re-match.
5.  Combine both sets of matches and aggregate to one row per
    `gvkey * cycle` with employee ideology statistics.

The chunk is set to `eval=FALSE` because processing the raw ContribDB
files takes significant time and disk space. The outputs
(`employee_panel_2000.rdata` -- `employee_panel_2024.rdata`) are already
saved.

```{r cycle-processing, eval=FALSE}
## ── Set to the ContribDB file we want to process :"contribDB_2018.csv", "contribDB_2020.csv", etc.
filename <- "contribDB_2018.csv"  # <-- CHANGE for each cycle
if (!file.exists(filename)) stop(paste("File not found:", filename,
  "\nMake sure the raw ContribDB CSV is in your working directory."))

cat("=== Processing:", filename, "===\n")
cat("File size:", round(file.size(filename) / 1e9, 1), "GB\n")

## ── Load the employer-name-to-gvkey lookup table ──
## all_exact: data.table with columns (employer_clean, gvkey, conm)
## Built in Section 3; contains 3,862 employer-name -> Compustat-gvkey pairs
load("employer_lookup_for_panel.rdata")  # loads `all_exact`
setDT(all_exact)
cat("Employer lookup:", nrow(all_exact), "rows\n")

## ── Suffix-cleaning function (must match the lookup construction exactly) ──
## Strips common corporate suffixes from uppercase employer strings
clean_employer <- function(x) {
  suffixes <- c("\\s+INC\\.?$", "\\s+INCORPORATED$", "\\s+CORP\\.?$",
                "\\s+CORPORATION$", "\\s+LLC$", "\\s+LLP$", "\\s+LP$",
                "\\s+LTD\\.?$", "\\s+LIMITED$", "\\s+CO\\.?$",
                "\\s+COMPANY$", "\\s+GROUP$", "\\s+HOLDINGS?$",
                "\\s+ENTERPRISES?$", "\\s+INTERNATIONAL$",
                "\\s+& CO\\.?$", "\\s+AND CO\\.?$",
                ",\\s*INC\\.?$", ",\\s*LLC$", ",\\s*LTD\\.?$",
                ",\\s*CORP\\.?$")
  for (suf in suffixes) x <- gsub(suf, "", x)
  gsub("\\s+", " ", trimws(x))
}

## ── Junk employer strings to exclude (self-employed, retired, etc.) ──
junk <- c("SELF", "SELF-EMPLOYED", "SELF EMPLOYED", "SELFEMPLOYED",
          "RETIRED", "NONE", "N/A", "NA", "NOT EMPLOYED",
          "HOMEMAKER", "HOME MAKER", "STUDENT", "UNEMPLOYED",
          "NOT APPLICABLE", "REFUSED", "INFORMATION REQUESTED",
          "INFORMATION REQUESTED PER BEST EFFORTS")

## ── Read only the 5 columns we need (keeps RAM ~2-4 GB instead of 27 GB) ──
cols_needed <- c("cycle", "bonica.cid", "contributor.type",
                 "contributor.employer", "contributor.cfscore")
cat("Reading file...\n")
t0 <- proc.time()
dt <- fread(filename, select = cols_needed, showProgress = TRUE)
elapsed_read <- round((proc.time() - t0)[3], 1)
cat("Read", format(nrow(dt), big.mark = ","), "rows in", elapsed_read, "seconds\n")

## Extract the election-cycle year from the first row
cycle_val <- dt[1, cycle]
cat("Cycle:", cycle_val, "\n")

## ── Filter: keep only individual contributors with valid cfscore and employer ──
dt <- dt[contributor.type == "I"]                          # individuals only
cat("Individuals:", format(nrow(dt), big.mark = ","), "\n")

dt[, contributor.cfscore := as.numeric(contributor.cfscore)]
dt <- dt[!is.na(contributor.cfscore)]                      # must have ideology score
cat("With cfscore:", format(nrow(dt), big.mark = ","), "\n")

dt <- dt[!is.na(contributor.employer) & contributor.employer != ""]
cat("With employer:", format(nrow(dt), big.mark = ","), "\n")

## ══════════════════════════════════════════════════════════════
## STEP 1: Fast pre-match (toupper + trimws only, NO suffix removal)
## This avoids running expensive regex on 50M+ rows
## ══════════════════════════════════════════════════════════════
cat("Step 1: Fast pre-match (toupper + trimws only)...\n")
t1 <- proc.time()
dt[, employer_raw := toupper(trimws(contributor.employer))]

## Remove junk employers
dt <- dt[!employer_raw %in% junk]
cat("After junk removal:", format(nrow(dt), big.mark = ","), "\n")

## Direct merge on the raw uppercase name against the lookup
matched1 <- merge(dt, all_exact[, .(employer_clean, gvkey)],
                  by.x = "employer_raw", by.y = "employer_clean")
cat("Step 1 matched rows:", format(nrow(matched1), big.mark = ","),
    "| gvkeys:", uniqueN(matched1$gvkey),
    "| time:", round((proc.time() - t1)[3], 1), "s\n")

## ══════════════════════════════════════════════════════════════
## STEP 2: Suffix-clean ONLY the unmatched rows (much smaller set)
## ══════════════════════════════════════════════════════════════
matched_raw <- unique(matched1$employer_raw)
unmatched <- dt[!employer_raw %in% matched_raw]
cat("Unmatched rows for suffix cleaning:", format(nrow(unmatched), big.mark = ","), "\n")

rm(dt); gc()  # free memory before the regex pass

t2 <- proc.time()
unmatched[, employer_clean := clean_employer(employer_raw)]

## Re-match the cleaned names against the lookup
matched2 <- merge(unmatched, all_exact[, .(employer_clean, gvkey)],
                  by = "employer_clean")
cat("Step 2 matched rows:", format(nrow(matched2), big.mark = ","),
    "| new gvkeys:", uniqueN(matched2$gvkey),
    "| time:", round((proc.time() - t2)[3], 1), "s\n")

rm(unmatched); gc()

## ── Combine matches from both steps ──
matched1[, employer_clean := employer_raw]
combined <- rbindlist(list(
  matched1[, .(cycle, bonica.cid, contributor.cfscore, gvkey)],
  matched2[, .(cycle, bonica.cid, contributor.cfscore, gvkey)]
))

rm(matched1, matched2); gc()

cat("Total matched rows:", format(nrow(combined), big.mark = ","), "\n")
cat("Unique gvkeys:", uniqueN(combined$gvkey), "\n")
cat("Unique donors:", format(uniqueN(combined$bonica.cid), big.mark = ","), "\n")

## ══════════════════════════════════════════════════════════════
## Aggregate to one row per gvkey * cycle
## ══════════════════════════════════════════════════════════════
cat("Aggregating...\n")
emp_panel <- combined[, .(
  emp_mean_cfscore   = mean(contributor.cfscore, na.rm = TRUE),   # mean employee ideology
  emp_median_cfscore = median(contributor.cfscore, na.rm = TRUE), # median employee ideology
  emp_sd_cfscore     = sd(contributor.cfscore, na.rm = TRUE),     # within-firm ideological dispersion
  n_emp_donors       = uniqueN(bonica.cid),                       # unique employee donors
  n_contributions    = .N                                          # total donation records
), by = .(gvkey, cycle)]

rm(combined); gc()

## ── Save the panel for this cycle ──
outfile <- paste0("employee_panel_", cycle_val, ".rdata")
save(emp_panel, file = outfile)
cat("\nSaved:", outfile, "\n")

## ── Print summary diagnostics ──
cat("\n=== Summary for cycle", cycle_val, "===\n")
cat("Firms (gvkey):", nrow(emp_panel), "\n")
cat("\nEmployee mean cfscore:\n"); print(summary(emp_panel$emp_mean_cfscore))
cat("\nDonor count:\n"); print(summary(emp_panel$n_emp_donors))
cat("\nFirms with >= 20 donors:", sum(emp_panel$n_emp_donors >= 20), "\n")
cat("Firms with >= 50 donors:", sum(emp_panel$n_emp_donors >= 50), "\n")
cat("Firms with >= 100 donors:", sum(emp_panel$n_emp_donors >= 100), "\n")

total_time <- round((proc.time() - t0)[3], 1)
cat("\nTotal time:", total_time, "seconds\n")
cat("Done!\n")
```

\newpage

# Section 2: WRDS Downloads -- Compustat and ExecuComp {#sec-wrds}

These chunks download the two WRDS datasets used in the project.

## 2.1 Compustat Fundamentals

Downloads annual firm financials from `comp.funda`, computes all outcome
variables (restructuring charges, layoffs, employee change) and lagged
controls, winsorizes continuous variables at the 1st/99th percentiles,
and saves as `compustat_outcomes.rdata`.

```{r wrds-compustat, eval=FALSE}
## ── Connect to WRDS via RPostgres ──
library(RPostgres)
wrds <- dbConnect(Postgres(),
                  host = "wrds-pgdata.wharton.upenn.edu", port = 9737,
                  dbname = "wrds", sslmode = "require",
                  user = Sys.getenv("WRDS_USER"),
                  password = Sys.getenv("WRDS_PW"))

## ── Query: annual Compustat fundamentals, fyear >= 1999, positive assets ──
## Restricts to industrial/standard/domestic/Canadian formats
compustat <- as.data.table(dbGetQuery(wrds, "
  SELECT gvkey, datadate, fyear, sich,
         at, sale, revt, cogs, xsga, ebitda, ni, ib, oibdp,
         spi, rcp, dp, xrd, capx, aqc, emp, csho, prcc_f,
         dltt, dlc, ceq, txdb, itcb, pstkl, che, act, lct,
         ppent, intan, gdwl, dvt
  FROM comp.funda
  WHERE fyear >= 1999
    AND at > 0
    AND indfmt = 'INDL' AND datafmt = 'STD'
    AND popsrc = 'D' AND consol = 'C'
"))

dbDisconnect(wrds)
setDT(compustat)
cat("Compustat raw:", nrow(compustat), "rows,", uniqueN(compustat$gvkey), "firms\n")

## ── Compute outcome and control variables ──
## SIC-2 industry code (first 2 digits of historical SIC)
compustat[, sic2 := floor(sich / 100)]

## Exclude financials (SIC 60-69) and utilities (SIC 49)
compustat[, exclude_industry := as.integer(sic2 %in% c(49, 60:69))]

## PRIMARY DV: neg_spi_at = -SPI/AT  (positive = more restructuring charges)
compustat[, neg_spi_at := -spi / at]

## Employee-based outcomes
compustat[, emp_change := (emp - shift(emp)) / shift(emp), by = gvkey]
compustat[, layoff := as.integer(emp_change < -0.05)]

## Restructuring charges / assets (alternative DV, low coverage)
compustat[, rcp_at := rcp / at]

## ── Control variables ──
compustat[, size := log(at)]
compustat[, roa := ib / at]
compustat[, leverage := (dltt + dlc) / at]
compustat[, tobinq := (at - ceq + csho * prcc_f) / at]
compustat[, sale_growth := (sale - shift(sale)) / shift(sale), by = gvkey]
compustat[, loss := as.integer(ni < 0)]
compustat[, capx_at := capx / at]
compustat[, rd_at := fifelse(is.na(xrd), 0, xrd / at)]
compustat[, acquisition := as.integer(!is.na(aqc) & aqc > 0)]
compustat[, emp_sale := emp / (sale / 1000)]  # employees per $M sales
compustat[, sga_at := xsga / at]
compustat[, cash_at := che / at]
compustat[, oper_margin := oibdp / sale]
compustat[oper_margin < -10, oper_margin := -10]  # cap extreme negatives
compustat[, asset_growth := (at - shift(at)) / shift(at), by = gvkey]

## Fix Inf values in growth variables
compustat[is.infinite(sale_growth), sale_growth := NA]
compustat[is.infinite(asset_growth), asset_growth := NA]

## NBER recession indicator (fiscal years during NBER contractions)
compustat[, nber_recession := as.integer(fyear %in% c(2001, 2008, 2009, 2020))]

## ── Lag all controls by one year ──
lag_vars <- c("size", "roa", "leverage", "tobinq", "sale_growth", "loss",
              "capx_at", "rd_at", "acquisition", "emp_sale", "sga_at", "cash_at",
              "oper_margin", "asset_growth", "neg_spi_at", "emp_change")
setorder(compustat, gvkey, fyear)
for (v in lag_vars) {
  compustat[, paste0("lag_", v) := shift(get(v), 1), by = gvkey]
}

## ── Winsorize continuous variables at 1st / 99th percentiles ──
winsorize <- function(x, probs = c(0.01, 0.99)) {
  q <- quantile(x, probs, na.rm = TRUE)
  x[x < q[1]] <- q[1]
  x[x > q[2]] <- q[2]
  x
}

win_vars <- c("neg_spi_at", "emp_change", "rcp_at",
              "size", "roa", "leverage", "tobinq", "sale_growth",
              "capx_at", "rd_at", "emp_sale", "sga_at", "cash_at", "oper_margin",
              grep("^lag_", names(compustat), value = TRUE))
for (v in intersect(win_vars, names(compustat))) {
  if (is.numeric(compustat[[v]])) compustat[, (v) := winsorize(get(v))]
}

save(compustat, file = "compustat_outcomes.rdata")
cat("Saved compustat_outcomes.rdata:", nrow(compustat), "rows,", ncol(compustat), "cols\n")
```

## 2.2 ExecuComp CEO Identification

Downloads CEO identity (one row per CEO-year) from ExecuComp's annual
compensation table, filtering on `ceoann = 'CEO'` and `year >= 2000`.

```{r wrds-execucomp, eval=FALSE}
library(RPostgres)
wrds <- dbConnect(Postgres(),
                  host = "wrds-pgdata.wharton.upenn.edu", port = 9737,
                  dbname = "wrds", sslmode = "require",
                  user = Sys.getenv("WRDS_USER"),
                  password = Sys.getenv("WRDS_PW"))

## Query: one row per CEO per firm per year
execucomp_ceos <- as.data.table(dbGetQuery(wrds, "
  SELECT gvkey, year, execid, exec_fullname
  FROM comp_execucomp.anncomp
  WHERE ceoann = 'CEO'
    AND year >= 2000
"))

dbDisconnect(wrds)
setDT(execucomp_ceos)
cat("ExecuComp CEOs:", nrow(execucomp_ceos), "rows\n")
cat("Unique gvkeys:", uniqueN(execucomp_ceos$gvkey), "\n")
cat("Unique execids:", uniqueN(execucomp_ceos$execid), "\n")

save(execucomp_ceos, file = "execucomp_ceos.rdata")
```

\newpage

# Section 3: Employer Lookup Construction {#sec-employer}

Builds the employer-name-to-Compustat-gvkey lookup table used by the
cycle processor (Section 1). This maps the free-text
`contributor.employer` field in DIME to standardized Compustat firm
identifiers.

The construction uses the full DIME contributor file
(`dime_contributors_1979_2024.rdata`, 44 M rows).

```{r employer-lookup, eval=FALSE}
## ── Load full DIME contributor file ──
library(data.table)
load("dime_contributors_1979_2024.rdata")  # contribs (44.2M rows)
setDT(contribs)

## ── Filter to relevant individuals ──
## Keep individuals with: last cycle >= 2000, non-missing cfscore, >= 2 distinct contributions
contribs <- contribs[contributor.type == "I" &
                       last_cycle_active >= 2000 &
                       !is.na(contributor.cfscore) &
                       num.distinct >= 2]
cat("After filter:", nrow(contribs), "rows\n")

## ── Standardize employer names ──
contribs[, employer_clean := toupper(trimws(most.recent.contributor.employer))]

## Remove junk employers (self-employed, retired, etc.)
junk <- c("SELF", "SELF-EMPLOYED", "SELF EMPLOYED", "SELFEMPLOYED",
          "RETIRED", "NONE", "N/A", "NA", "NOT EMPLOYED",
          "HOMEMAKER", "HOME MAKER", "STUDENT", "UNEMPLOYED")
contribs <- contribs[!employer_clean %in% junk & employer_clean != ""]

## Remove corporate suffixes (same function as Section 1)
suffixes <- c("\\s+INC\\.?$", "\\s+INCORPORATED$", "\\s+CORP\\.?$",
              "\\s+CORPORATION$", "\\s+LLC$", "\\s+LLP$", "\\s+LP$",
              "\\s+LTD\\.?$", "\\s+LIMITED$", "\\s+CO\\.?$",
              "\\s+COMPANY$", "\\s+GROUP$", "\\s+HOLDINGS?$",
              "\\s+ENTERPRISES?$", "\\s+INTERNATIONAL$",
              "\\s+& CO\\.?$", "\\s+AND CO\\.?$",
              ",\\s*INC\\.?$", ",\\s*LLC$", ",\\s*LTD\\.?$",
              ",\\s*CORP\\.?$")
for (suf in suffixes) contribs[, employer_clean := gsub(suf, "", employer_clean)]
contribs[, employer_clean := gsub("\\s+", " ", trimws(employer_clean))]

## ── Count unique donors per employer, keep those with >= 10 ──
emp_counts <- contribs[, .(n_donors = uniqueN(bonica.cid)), by = employer_clean]
emp_counts <- emp_counts[n_donors >= 10]

## ── Match to Compustat firm names (exact + manual aliases) ──
## Load Compustat to get official company names
load("compustat_outcomes.rdata")
cs_names <- unique(compustat[, .(gvkey, conm)])
cs_names[, conm_clean := toupper(trimws(conm))]
for (suf in suffixes) cs_names[, conm_clean := gsub(suf, "", conm_clean)]
cs_names[, conm_clean := gsub("\\s+", " ", trimws(conm_clean))]

## Exact matches: employer_clean == conm_clean
exact_matches <- merge(emp_counts, cs_names[, .(employer_clean = conm_clean, gvkey, conm)],
                        by = "employer_clean")

## Manual aliases for known name changes / variants
manual_aliases <- data.table(
  employer_clean = c(
    "GOOGLE", "ALPHABET",
    "FACEBOOK", "META", "META PLATFORMS",
    "HEWLETT PACKARD", "HEWLETT-PACKARD", "HP",
    "JPMORGAN CHASE", "JP MORGAN", "JP MORGAN CHASE", "JPMORGAN",
    "CHASE MANHATTAN",
    "GOLDMAN SACHS",
    "MORGAN STANLEY",
    "BANK OF AMERICA",
    "MERRILL LYNCH",
    "BEAR STEARNS",
    "LEHMAN BROTHERS",
    "COUNTRYWIDE", "COUNTRYWIDE FINANCIAL",
    "CITIBANK", "CITICORP",
    "PHILIP MORRIS", "ALTRIA", "ALTRIA GROUP",
    "KRAFT FOODS", "KRAFT",
    "TIME WARNER",
    "WARNER BROS",
    "RAYTHEON", "RAYTHEON TECHNOLOGIES",
    "RTX",
    "WORLDCOM", "MCI", "MCI WORLDCOM",
    "LUCENT", "LUCENT TECHNOLOGIES",
    "TRIBUNE", "TRIBUNE MEDIA",
    "ANDERSEN CONSULTING", "ACCENTURE",
    "WAL MART", "WAL-MART",
    "SUNTRUST BANK", "SUNTRUST",
    "UNITEDHEALTH", "UNITED HEALTH GROUP", "UNITED HEALTHCARE",
    "ANHEUSER BUSCH", "ANHEUSER-BUSCH", "ANHEUSER BUSCH COS"
  ),
  gvkey = c(
    "160329", "160329",
    "170617", "170617", "170617",
    "005606", "005606", "005606",
    "002968", "002968", "002968", "002968",
    "002968",
    "114628",
    "012124",
    "007647",
    "007267",
    "011818",
    "030128",
    "003555", "003555",
    "003243", "003243",
    "008543", "008543", "008543",
    "012978", "012978",
    "025056",
    "164296",
    "008972", "010983",
    "010983",
    "143972", "143972", "143972",
    "062599", "062599",
    "010726", "010726",
    "143357", "143357",
    "011259", "011259",
    "010187", "010187",
    "010903", "010903", "010903",
    "001663", "001663", "001663"
  )
)

manual_aliases <- merge(manual_aliases, cs_names[, .(gvkey, conm)], by = "gvkey", all.x = TRUE)

## Combine exact + manual into final lookup
all_exact <- rbindlist(list(
  exact_matches[, .(employer_clean, gvkey, conm)],
  manual_aliases[, .(employer_clean, gvkey, conm)]
), fill = TRUE)
all_exact <- unique(all_exact, by = c("employer_clean", "gvkey"))

cat("Final employer lookup:", nrow(all_exact), "rows\n")
cat("Unique gvkeys:", uniqueN(all_exact$gvkey), "\n")

save(all_exact, file = "employer_lookup_for_panel.rdata")
cat("Saved: employer_lookup_for_panel.rdata\n")
```

\newpage

# Section 4: CEO DIME Matching {#sec-ceo-match}

Matches ExecuComp CEO names to DIME contributor records to obtain a
lifetime CFscore for each CEO. The procedure:

1.  Clean both name sets: uppercase, remove punctuation, strip suffixes
    (Jr, Sr, II, III, MD, PhD, MBA, etc.).
2.  Parse to `LASTNAME FIRSTNAME` format (DIME is "last, first";
    ExecuComp is "first last").
3.  Merge on the two-token name key via `data.table` join (fast --
    seconds not hours).
4.  Disambiguate multiple matches: prefer candidates whose DIME employer
    matches the Compustat firm, then break ties by `num.distinct` (more
    distinct contributions = more reliable score).

Result: `ceo_dime_lookup.rdata` with 4,157 of 8,180 ExecuComp CEOs
matched (50.8%).

```{r ceo-match, eval=FALSE}
library(data.table)
setwd("/Users/homemac/Library/Mobile Documents/com~apple~CloudDocs/My Research/Accounting Research/Political_Accounting/Political Misalignment/Codes")

## ── Name cleaning function ──
## Uppercases, removes periods/commas, strips personal suffixes
clean_name <- function(x) {
  x <- toupper(x)
  x <- gsub('[.,]', ' ', x)
  x <- gsub('\\b(JR|SR|II|III|IV|V|MD|PHD|MBA|CFA|CPA|ESQ|BSC|DDS|DO)\\b', '', x)
  x <- gsub('\\s+', ' ', trimws(x))
  x
}

## ══════════════════════════════════════════════════════════════
## Build DIME lookup table
## ══════════════════════════════════════════════════════════════
## dime_contribs_slim.rdata: pre-filtered DIME contributor table
## Columns: bonica.cid, most.recent.contributor.name, contributor.cfscore,
##          most.recent.contributor.employer, num.distinct
load("dime_contribs_slim.rdata")  # loads `contribs`
setDT(contribs)

dime_names <- contribs[, .(bonica.cid,
                           name = most.recent.contributor.name,
                           cfscore = contributor.cfscore,
                           employer = most.recent.contributor.employer,
                           num_distinct = num.distinct)]
rm(contribs); gc()

## Clean DIME names and create a two-token key: "LASTNAME FIRSTNAME"
## DIME format: "lastname, firstname [middle]" -> after clean: "LASTNAME FIRSTNAME MIDDLE"
## We keep only the first two tokens
dime_names[, name_clean := clean_name(name)]
dime_names[, name_key := gsub(',', '', name_clean)]       # remove leftover commas
dime_names[, name_key := gsub('\\s+', ' ', trimws(name_key))]
dime_names[, name_key := sub('^(\\S+\\s+\\S+).*', '\\1', name_key)]  # first 2 tokens

## ══════════════════════════════════════════════════════════════
## Clean ExecuComp names into the same "LASTNAME FIRSTNAME" format
## ══════════════════════════════════════════════════════════════
load("execucomp_ceos.rdata")
setDT(execucomp_ceos)
execucomp_ceos[, execid := as.character(execid)]
execucomp_ceos[, name_clean := clean_name(exec_fullname)]

## ExecuComp format: "FIRSTNAME [MIDDLE] LASTNAME" -> reverse to "LASTNAME FIRSTNAME"
name_parts <- strsplit(execucomp_ceos$name_clean, ' ')
execucomp_ceos[, name_key := sapply(name_parts, function(p) {
  if (length(p) >= 2) paste(p[length(p)], p[1]) else p[1]
})]

## ══════════════════════════════════════════════════════════════
## Merge ExecuComp x DIME on name_key (cartesian join)
## ══════════════════════════════════════════════════════════════
exec_unique <- unique(execucomp_ceos[, .(execid, name_key, exec_fullname)])

matches <- merge(execucomp_ceos[, .(execid, gvkey, year, name_key, exec_fullname)],
                 dime_names[, .(bonica.cid, name_key, cfscore, employer, num_distinct)],
                 by = 'name_key', allow.cartesian = TRUE)
cat("Raw matches:", format(nrow(matches), big.mark = ","), "\n")
cat("Unique execids matched:", uniqueN(matches$execid), "out of",
    uniqueN(exec_unique$execid), "\n")

## ══════════════════════════════════════════════════════════════
## Disambiguate: prefer DIME employer matching the Compustat firm
## ══════════════════════════════════════════════════════════════
load("employer_lookup_for_panel.rdata")  # all_exact

## Flag if the DIME contributor's employer name matches the Compustat firm
matches <- merge(matches, all_exact[, .(gvkey, employer_clean)],
                 by = 'gvkey', allow.cartesian = TRUE)
matches[, dime_employer_clean := toupper(trimws(employer))]
matches[, employer_match := !is.na(dime_employer_clean) & !is.na(employer_clean) &
          (grepl(employer_clean, dime_employer_clean, fixed = TRUE) |
           grepl(dime_employer_clean, employer_clean, fixed = TRUE))]

## Pick best match per execid: employer match first, then highest num_distinct
matches[, priority := ifelse(employer_match, 1L, 2L)]
setorder(matches, execid, priority, -num_distinct)
ceo_lookup <- unique(matches, by = 'execid')

cat("Final CEO lookup:", nrow(ceo_lookup), "unique execids\n")
cat("  With employer match:", sum(ceo_lookup$employer_match), "\n")
cat("  Without (name only):", sum(!ceo_lookup$employer_match), "\n")

## ── Save the slim lookup for use in panel assembly ──
ceo_dime_lookup <- ceo_lookup[, .(execid, exec_fullname, bonica.cid,
                                   ceo_cfscore = cfscore, num_distinct,
                                   name_key, employer_match,
                                   dime_employer = employer)]
save(ceo_dime_lookup, file = "ceo_dime_lookup.rdata")
cat("Saved: ceo_dime_lookup.rdata (", nrow(ceo_dime_lookup), "rows)\n")
```

\newpage

# Section 5: Panel Stacking and Assembly {#sec-panel}

Stacks all 13 cycle-level employee panels into a single firm-cycle
panel, merges CEO identity from ExecuComp (with year-1 fallback for
missing matches), attaches CEO CFscores from the DIME lookup, computes
misalignment, adds company names, and filters to a quality sample (\>=
20 employee donors).

```{r stack-and-merge, eval=FALSE}
library(data.table)
setwd("/Users/homemac/Library/Mobile Documents/com~apple~CloudDocs/My Research/Accounting Research/Political_Accounting/Political Misalignment/Codes")

## ══════════════════════════════════════════════════════════════
## STEP 1: Stack all 13 employee panels (2000-2024)
## ══════════════════════════════════════════════════════════════
files <- sort(list.files(pattern = "^employee_panel_20[0-9]{2}\\.rdata$"))
cat("Found", length(files), "panel files:", paste(files, collapse = ", "), "\n")

panels <- list()
for (f in files) {
  env <- new.env()
  load(f, envir = env)
  obj_name <- ls(envir = env)[1]
  panels[[f]] <- get(obj_name, envir = env)
}
panel <- rbindlist(panels, fill = TRUE)
rm(panels); gc()

cat("Stacked panel:", format(nrow(panel), big.mark = ","), "firm-cycles\n")
cat("Unique firms:", uniqueN(panel$gvkey), "\n")
cat("Cycles:", paste(sort(unique(panel$cycle)), collapse = ", "), "\n")

## ══════════════════════════════════════════════════════════════
## STEP 2: Merge CEO identity from ExecuComp
## ══════════════════════════════════════════════════════════════
load("execucomp_ceos.rdata")
setDT(execucomp_ceos)
ceo_slim <- unique(execucomp_ceos[, .(gvkey, year, execid, exec_fullname)])

## Primary merge: election cycle year == ExecuComp year
panel <- merge(panel, ceo_slim,
               by.x = c("gvkey", "cycle"),
               by.y = c("gvkey", "year"),
               all.x = TRUE)
cat("CEO identified (direct year match):", format(sum(!is.na(panel$execid)), big.mark=","),
    "of", format(nrow(panel), big.mark=","), "\n")

## Fallback: if no CEO in cycle year, try year - 1
missing_idx <- which(is.na(panel$execid))
cat("Missing CEO:", format(length(missing_idx), big.mark=","), "-- trying year-1 fallback\n")

if (length(missing_idx) > 0) {
  missing_dt <- panel[missing_idx, .(gvkey, cycle, row_idx = missing_idx)]
  missing_dt[, year_m1 := cycle - 1L]

  fallback <- merge(missing_dt, ceo_slim,
                    by.x = c("gvkey", "year_m1"),
                    by.y = c("gvkey", "year"),
                    all.x = TRUE)
  fallback <- fallback[!is.na(execid)]
  fallback <- unique(fallback, by = "row_idx")  # one CEO per firm-cycle

  if (nrow(fallback) > 0) {
    panel[fallback$row_idx, `:=`(execid = fallback$execid,
                                  exec_fullname = fallback$exec_fullname)]
    cat("Fallback matched:", format(nrow(fallback), big.mark=","), "\n")
  }
}
cat("Total CEO identified:", format(sum(!is.na(panel$execid)), big.mark=","),
    "of", format(nrow(panel), big.mark=","), "\n")

## ══════════════════════════════════════════════════════════════
## STEP 3: Merge CEO CFscore from DIME lookup
## ══════════════════════════════════════════════════════════════
load("ceo_dime_lookup.rdata")
setDT(ceo_dime_lookup)
ceo_cf <- unique(ceo_dime_lookup[, .(execid, ceo_cfscore)], by = "execid")
ceo_cf <- ceo_cf[!is.na(ceo_cfscore)]

panel <- merge(panel, ceo_cf, by = "execid", all.x = TRUE)
cat("Firm-cycles with CEO CFscore:", format(sum(!is.na(panel$ceo_cfscore)), big.mark=","), "\n")

## ══════════════════════════════════════════════════════════════
## STEP 4: Compute misalignment measures
## ══════════════════════════════════════════════════════════════
## misalign_abs:    absolute difference between CEO and employee mean ideology
## misalign_signed: CEO minus employee mean (positive = CEO more conservative)
panel[, misalign_abs := abs(ceo_cfscore - emp_mean_cfscore)]
panel[, misalign_signed := ceo_cfscore - emp_mean_cfscore]

## ══════════════════════════════════════════════════════════════
## STEP 5: Add company names from employer lookup
## ══════════════════════════════════════════════════════════════
load("employer_lookup_for_panel.rdata")
setDT(all_exact)
firm_names <- unique(all_exact[, .(gvkey, conm)], by = "gvkey")
panel <- merge(panel, firm_names, by = "gvkey", all.x = TRUE)

## ══════════════════════════════════════════════════════════════
## STEP 6: Create panel versions and save
## ══════════════════════════════════════════════════════════════
panel_complete <- panel[!is.na(misalign_abs)]       # non-missing misalignment
panel_qual     <- panel_complete[n_emp_donors >= 20] # quality filter: >= 20 donors

cat("\n--- Pipeline funnel ---\n")
cat("Total firm-cycles (stacked):", format(nrow(panel), big.mark=","), "\n")
cat("Computable misalignment:", format(nrow(panel_complete), big.mark=","), "\n")
cat("Quality filtered (>=20 donors):", format(nrow(panel_qual), big.mark=","), "\n")
cat("Unique firms (quality):", uniqueN(panel_qual$gvkey), "\n")

save(panel, file = "panel_full.rdata")
save(panel_complete, file = "panel_complete.rdata")
save(panel_qual, file = "panel_qual.rdata")
cat("Saved panel_full.rdata, panel_complete.rdata, panel_qual.rdata\n")
```

\newpage

# Section 6: Regression Data Construction {#sec-regdata}

Merges the quality-filtered misalignment panel with Compustat outcomes,
then constructs all additional variables needed for the regressions.

**Bug fixes integrated here** (from `fix_dupes4.R`, `fix_turnover2.R`,
`fix_firstyear.R`):

-   **Deduplication:** After the Compustat merge, some firm-years have
    multiple CEO rows. We keep one row per `gvkey * fyear_lead`,
    preferring the row with a non-missing `ceo_cfscore`.
-   **CEO turnover:** Computed at the firm-year level by comparing the
    set of CEO `execid`s across consecutive ExecuComp years (not
    row-level shifts, which produced incorrect values when multiple CEOs
    appear).
-   **First-year CEO:** Defined as the first panel appearance of a given
    `execid` at a given `gvkey` (by cycle order), not the ExecuComp
    first year.

```{r build-reg-data, eval=FALSE}
library(data.table)
setwd("/Users/homemac/Desktop")

## ══════════════════════════════════════════════════════════════
## PART A: Merge quality panel with Compustat
## ══════════════════════════════════════════════════════════════
load("panel_qual.rdata")
load("compustat_outcomes.rdata")
setDT(panel_qual); setDT(compustat)

## Lagged specification: misalignment at cycle t -> outcomes at fiscal year t+1
panel_qual[, fyear_lead := cycle + 1L]

reg_data <- merge(panel_qual, compustat,
                  by.x = c("gvkey", "fyear_lead"),
                  by.y = c("gvkey", "fyear"),
                  all.x = TRUE)

## Drop rows without Compustat match
reg_data <- reg_data[!is.na(at)]
cat("After Compustat merge:", nrow(reg_data), "rows\n")

## Exclude financials (SIC 60-69) and utilities (SIC 49)
reg_data <- reg_data[exclude_industry == 0 | is.na(exclude_industry)]
reg_data <- reg_data[!is.na(neg_spi_at) | !is.na(emp_change)]  # need at least one DV
cat("After industry exclusion:", nrow(reg_data), "rows\n")

## ══════════════════════════════════════════════════════════════
## PART B: Deduplicate (fix from fix_dupes4.R)
## Some firm-years map to multiple CEO rows from ExecuComp.
## Keep one row per gvkey * fyear_lead, preferring non-missing ceo_cfscore.
## ══════════════════════════════════════════════════════════════
reg_data[, has_cfs := as.integer(!is.na(ceo_cfscore))]
setorder(reg_data, gvkey, fyear_lead, -has_cfs)
reg_data <- reg_data[!duplicated(reg_data[, .(gvkey, fyear_lead)])]
reg_data[, has_cfs := NULL]
cat("After dedup:", nrow(reg_data), "rows\n")
cat("gvkey+fyear dupes:", sum(duplicated(reg_data[, .(gvkey, fyear_lead)])), "\n")

## ══════════════════════════════════════════════════════════════
## PART C: Construct additional variables (from build_reg_data_v2.R)
## ══════════════════════════════════════════════════════════════

## ── Quadrant indicators ──
## Classify each firm-cycle into one of four ideological configurations
reg_data[, cross_ceo_right := as.integer(ceo_cfscore > 0 & emp_mean_cfscore < 0)]
reg_data[, cross_ceo_left  := as.integer(ceo_cfscore < 0 & emp_mean_cfscore > 0)]
reg_data[, same_side_right := as.integer(ceo_cfscore > emp_mean_cfscore &
                                          cross_ceo_right == 0 & cross_ceo_left == 0)]
reg_data[, same_side_left  := as.integer(ceo_cfscore < emp_mean_cfscore &
                                          cross_ceo_right == 0 & cross_ceo_left == 0)]

## Quadrant label (for diagnostics)
reg_data[, quadrant := fcase(
  cross_ceo_right == 1, "Cross: CEO Right / Emp Left",
  cross_ceo_left == 1,  "Cross: CEO Left / Emp Right",
  same_side_right == 1, "Same side: CEO more Right",
  same_side_left == 1,  "Same side: CEO more Left",
  default = "Exact tie or NA"
)]

## ── CEO absolute conservative indicator ──
reg_data[, ceo_abs_conservative := as.integer(ceo_cfscore > 0)]

## ── Log number of donors ──
reg_data[, log_n_donors := log(n_emp_donors)]

## ── SIC 2-digit industry code ──
reg_data[, sic2 := floor(sich / 100)]
## Fill missing sic2 from Compustat if needed
if (sum(is.na(reg_data$sic2)) > 0) {
  cs_sic <- compustat[!is.na(sich), .(sich_cs = floor(sich[.N] / 100)), by = gvkey]
  reg_data <- merge(reg_data, cs_sic, by = "gvkey", all.x = TRUE)
  reg_data[is.na(sic2) & !is.na(sich_cs), sic2 := sich_cs]
  reg_data[, sich_cs := NULL]
}

## ── Donor coverage (employee donors / total employees in thousands) ──
if ("emp" %in% names(reg_data)) {
  reg_data[, donor_coverage := fifelse(is.na(emp) | emp == 0, NA_real_,
                                        n_emp_donors / (emp * 1000))]
}

## ── Lagged misalignment from prior election cycle ──
## Cycles are biennial, so shift(1) by gvkey gives the prior cycle (2 years back)
setorder(reg_data, gvkey, cycle)
reg_data[, misalign_abs_L2 := shift(misalign_abs, 1), by = gvkey]

## ══════════════════════════════════════════════════════════════
## PART D: CEO Turnover (fix from fix_turnover2.R)
## Compute at firm-year level by comparing CEO sets across consecutive years
## ══════════════════════════════════════════════════════════════
load("execucomp_ceos.rdata")
ec <- as.data.table(execucomp_ceos)
ec[, execid := as.character(execid)]

## For each gvkey-year, collect all CEO execids as a sorted string
ceo_set <- ec[, .(ceo_ids = paste(sort(unique(execid)), collapse = ",")),
              by = .(gvkey, year)]
setorder(ceo_set, gvkey, year)
ceo_set[, prev_ceo_ids := shift(ceo_ids, 1), by = gvkey]
ceo_set[, prev_year := shift(year, 1), by = gvkey]

## Turnover = CEO set changed from prior consecutive year
ceo_set[, ceo_turnover := as.integer(prev_year == year - 1 & ceo_ids != prev_ceo_ids)]
ceo_set[prev_year != year - 1 | is.na(prev_year), ceo_turnover := NA_integer_]

turnover_dt <- ceo_set[, .(gvkey, fyear_lead = year, ceo_turnover)]
reg_data[, execid := as.character(execid)]
reg_data <- merge(reg_data, turnover_dt, by = c("gvkey", "fyear_lead"), all.x = TRUE)

## ══════════════════════════════════════════════════════════════
## PART E: First-year CEO (fix from fix_firstyear.R)
## First panel appearance of this execid at this gvkey
## ══════════════════════════════════════════════════════════════
setorder(reg_data, gvkey, execid, cycle)
reg_data[, first_year_ceo := as.integer(cycle == min(cycle)), by = .(gvkey, execid)]

## ══════════════════════════════════════════════════════════════
## PART F: Industry downturn measures (from run_industry_downturn.R)
## ══════════════════════════════════════════════════════════════
cs <- copy(compustat)
cs <- cs[!sic2 %in% c(49, 60:69)]

## Industry-level aggregates by sic2 x fyear
ind <- cs[!is.na(sic2) & !is.na(fyear), .(
  ind_sale_growth  = median(sale_growth, na.rm = TRUE),
  ind_neg_spi      = median(neg_spi_at, na.rm = TRUE),
  ind_layoff_rate  = mean(emp_change < -0.05, na.rm = TRUE),  # fraction of firms with layoffs
  n_firms          = .N
), by = .(sic2, fyear)]

ind[, ind_downturn := as.integer(ind_sale_growth < 0)]
ind <- ind[!is.na(ind_sale_growth)]

## Merge to reg_data
reg_data <- merge(reg_data,
                  ind[, .(sic2, fyear_lead = fyear, ind_sale_growth,
                          ind_neg_spi, ind_downturn, ind_layoff_rate)],
                  by = c("sic2", "fyear_lead"), all.x = TRUE)
rm(cs, ind); gc()

## ══════════════════════════════════════════════════════════════
## Save final regression data
## ══════════════════════════════════════════════════════════════
cat("\n=== FINAL reg_data ===\n")
cat("Dimensions:", nrow(reg_data), "rows x", ncol(reg_data), "cols\n")
cat("Unique firms:", uniqueN(reg_data$gvkey), "\n")
cat("gvkey+fyear dupes:", sum(duplicated(reg_data[, .(gvkey, fyear_lead)])), "\n")
cat("CEO turnover:\n"); print(table(reg_data$ceo_turnover, useNA = "ifany"))
cat("First year CEO:\n"); print(table(reg_data$first_year_ceo, useNA = "ifany"))

save(reg_data, file = "reg_data_v3.rdata")
cat("Saved reg_data_v3.rdata\n")
```

\newpage

# Section 7: Load Processed Data {#sec-load}

This is the entry point for running the analysis tables below. It loads
the final regression dataset and prepares the variables used across all
models.

```{r load-data}
## ── Load the final regression dataset ──
load("reg_data_v3.rdata")
cat("reg_data loaded:", nrow(reg_data), "rows x", ncol(reg_data), "cols\n")

## ── Rename fyear_lead -> fyear for cleaner FE notation ──
if (!"fyear" %in% names(reg_data)) setnames(reg_data, "fyear_lead", "fyear")

## ── Prepare the analysis sample ──
## Exclude financials (SIC 60-69) and utilities (SIC 49)
dt <- reg_data[!sic2 %in% c(49, 60:69)]
cat("Analysis sample (excl. fin/util):", nrow(dt), "rows,",
    uniqueN(dt$gvkey), "firms\n")

## ── Create derived variables used across multiple tables ──
## recession: NBER recession indicator (fiscal years 2001, 2008-09, 2020)
dt[, recession := nber_recession]

## lag_aqc_indicator: rename for clarity in formulas
dt[, lag_aqc_indicator := lag_acquisition]

## Asymmetric decomposition of signed misalignment
## ceo_right = max(signed_misalign, 0): amount by which CEO is more conservative
## ceo_left  = max(-signed_misalign, 0): amount by which CEO is more liberal
dt[, ceo_right := pmax(misalign_signed, 0)]
dt[, ceo_left  := pmax(-misalign_signed, 0)]

## ── Standard control variable string (used in all model formulas) ──
ctrl <- paste("lag_size + lag_roa + lag_leverage + lag_tobinq +",
              "lag_sale_growth + lag_loss + lag_capx_at + lag_rd_at +",
              "lag_aqc_indicator")
cat("Controls:", ctrl, "\n")

## ── Helper functions for table construction ──
## Format numbers for table cells
fmt  <- function(x, d = 3) formatC(x, format = "f", digits = d, big.mark = ",")
fmt0 <- function(x) formatC(x, format = "d", big.mark = ",")
fmt4 <- function(x) formatC(x, format = "f", digits = 4)

## Significance stars
stars <- function(p) {
  ifelse(is.na(p), "",
    ifelse(p < 0.01, "***",
      ifelse(p < 0.05, "**",
        ifelse(p < 0.10, "*", ""))))
}

## Extract coefficient, SE, and p-value for a variable from a fixest model
get_cp <- function(mod, var) {
  b <- coef(mod); V <- vcov(mod)
  idx <- match(var, names(b))
  if (is.na(idx)) return(c(NA_real_, NA_real_, NA_real_))
  se <- sqrt(V[idx, idx])
  p <- 2 * pt(abs(b[idx] / se), df = degrees_freedom(mod, "t"), lower.tail = FALSE)
  unname(c(b[idx], se, p))
}

## Compute the net effect of two coefficients (e.g., base + interaction * scale)
net_eff <- function(mod, v1, v2, scale = 1) {
  b <- coef(mod); V <- vcov(mod)
  i1 <- match(v1, names(b)); i2 <- match(v2, names(b))
  if (is.na(i1) || is.na(i2)) return(c(NA_real_, NA_real_, NA_real_))
  net <- b[i1] + b[i2] * scale
  se <- sqrt(V[i1, i1] + scale^2 * V[i2, i2] + 2 * scale * V[i1, i2])
  p <- 2 * pt(abs(net / se), df = degrees_freedom(mod, "t"), lower.tail = FALSE)
  unname(c(net, se, p))
}

## Get within-R-squared (or overall R2 if within unavailable)
get_r2 <- function(mod) {
  wr2 <- fitstat(mod, "wr2")[[1]]
  if (is.na(wr2)) return(fitstat(mod, "r2")[[1]])
  return(wr2)
}

## Build a two-row block (coefficient + SE) for a variable across a list of models
make_row <- function(var, label, mods) {
  coefs <- sapply(mods, function(m) {
    r <- get_cp(m, var)
    if (is.na(r[1])) return(c("", ""))
    c(paste0(fmt4(r[1]), stars(r[3])), paste0("(", fmt4(r[2]), ")"))
  })
  if (!is.matrix(coefs)) coefs <- matrix(coefs, nrow = 2, ncol = length(mods))
  row1 <- c(label, coefs[1, ])
  row2 <- c("", coefs[2, ])
  rbind(row1, row2)
}
```

\newpage

# Section 8: Paper Tables {#sec-tables}

All seven final tables for the paper, estimated from the loaded
`reg_data_v3`.

## Table 1: Sample Construction

Shows the step-by-step construction of the analysis sample from 13
stacked cycle panels down to the final regression sample.

```{r table1}
## Load the intermediate panel objects for the funnel counts
load("panel_full.rdata")
pf <- panel
load("panel_complete.rdata")
pc <- panel_complete
load("panel_qual.rdata")
pq <- panel_qual

dt_spi <- dt[!is.na(neg_spi_at)]

samp <- data.frame(
  Step = 1:6,
  Description = c(
    "All 13 cycle panels stacked (2000--2024)",
    "Non-missing CEO and employee ideology",
    "At least 20 employee donors per firm-cycle",
    "Merged with Compustat outcomes",
    "Excl. financials (SIC 60--67) and utilities (SIC 49)",
    "Non-missing restructuring charges DV"
  ),
  FC = c(fmt0(nrow(pf)), fmt0(nrow(pc)), fmt0(nrow(pq)),
         fmt0(nrow(reg_data)), fmt0(nrow(dt)), fmt0(nrow(dt_spi))),
  UF = c(fmt0(uniqueN(pf$gvkey)), fmt0(uniqueN(pc$gvkey)),
         fmt0(uniqueN(pq$gvkey)), fmt0(uniqueN(reg_data$gvkey)),
         fmt0(uniqueN(dt$gvkey)), fmt0(uniqueN(dt_spi$gvkey))),
  stringsAsFactors = FALSE
)

kbl(samp, booktabs = TRUE, align = c("c","l","r","r"),
    caption = "Sample Construction",
    col.names = c("Step","Description","Firm-Cycles","Unique Firms")) %>%
  kable_styling(latex_options = c("hold_position"), font_size = 9,
                full_width = FALSE, position = "center") %>%
  footnote(
    general = "The sample is a biennial panel covering 13 election cycles (2000--2024). CEO ideology is a lifetime DIME CFscore matched from ExecuComp to DIME contributor records. Employee ideology is the mean CFscore of employee political donors per firm-cycle. Financials = SIC 6000--6799; Utilities = SIC 4900--4999. Restructuring charges = -SPI/AT (positive values indicate larger charges).",
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )

rm(pf, pc, pq, panel, panel_complete, panel_qual)
```

\newpage

## Table 2: Descriptive Statistics

Summary statistics for ideology measures, outcome variables, moderators,
and lagged controls.

```{r table2}
## Helper: one row of descriptive statistics
desc_row <- function(label, x) {
  data.frame(
    Variable = label,
    N = fmt0(sum(!is.na(x))),
    Mean = fmt(mean(x, na.rm = TRUE)),
    SD = fmt(sd(x, na.rm = TRUE)),
    P25 = fmt(as.numeric(quantile(x, 0.25, na.rm = TRUE))),
    Median = fmt(median(x, na.rm = TRUE)),
    P75 = fmt(as.numeric(quantile(x, 0.75, na.rm = TRUE))),
    stringsAsFactors = FALSE
  )
}

## Panel A: Ideology and Misalignment
panA <- rbind(
  desc_row("|Misalignment|", dt$misalign_abs),
  desc_row("Signed Misalignment", dt$misalign_signed),
  desc_row("CEO CFscore", dt$ceo_cfscore),
  desc_row("Employee Mean CFscore", dt$emp_mean_cfscore),
  desc_row("Employee CFscore SD", dt$emp_sd_cfscore),
  desc_row("No. Employee Donors", dt$n_emp_donors),
  desc_row("Donor Coverage", dt$donor_coverage)
)

## Panel B: Outcome Variables
panB <- rbind(
  desc_row("Restructuring Charges (neg SPI/AT)", dt$neg_spi_at),
  desc_row("Layoff Indicator", dt$layoff),
  desc_row("Employee Change", dt$emp_change)
)

## Panel C: Moderators
panC <- rbind(
  desc_row("NBER Recession", dt$recession),
  desc_row("Industry Peer Layoff Rate", dt$ind_layoff_rate),
  desc_row("Industry Downturn", dt$ind_downturn)
)

## Panel D: Lagged Controls
panD <- rbind(
  desc_row("Firm Size (log assets)", dt$lag_size),
  desc_row("ROA", dt$lag_roa),
  desc_row("Leverage", dt$lag_leverage),
  desc_row("Tobin's Q", dt$lag_tobinq),
  desc_row("Sales Growth", dt$lag_sale_growth),
  desc_row("Loss Indicator", dt$lag_loss),
  desc_row("Capex / Assets", dt$lag_capx_at),
  desc_row("R&D / Assets", dt$lag_rd_at),
  desc_row("Acquisition Indicator", dt$lag_aqc_indicator)
)

all_desc <- rbind(panA, panB, panC, panD)

kbl(all_desc, booktabs = TRUE, align = c("l", rep("r", 6)),
    caption = "Descriptive Statistics",
    col.names = c("Variable","N","Mean","SD","P25","Median","P75")) %>%
  kable_styling(latex_options = c("hold_position"), font_size = 9,
                full_width = FALSE, position = "center") %>%
  pack_rows("Panel A: Ideology and Misalignment", 1, nrow(panA), bold = TRUE) %>%
  pack_rows("Panel B: Outcomes", nrow(panA)+1, nrow(panA)+nrow(panB), bold = TRUE) %>%
  pack_rows("Panel C: Moderators", nrow(panA)+nrow(panB)+1,
            nrow(panA)+nrow(panB)+nrow(panC), bold = TRUE) %>%
  pack_rows("Panel D: Lagged Controls", nrow(panA)+nrow(panB)+nrow(panC)+1,
            nrow(all_desc), bold = TRUE) %>%
  footnote(
    general = paste0("Sample: ", fmt0(nrow(dt)), " firm-cycle observations (",
                     fmt0(uniqueN(dt$gvkey)),
                     " firms). Financials and utilities excluded. All controls are lagged one period. Restructuring charges = -SPI/AT. Layoff = 1 if employee change < -5 percent. Donor coverage = no. employee donors / total employees."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

\newpage

## Table 3: Baseline -- Misalignment and Restructuring Charges

Tests whether CEO-employee ideological misalignment predicts
restructuring charges, building up from pooled OLS to the full
specification with firm + year FE and additional controls.

```{r table3}
## (1) Pooled OLS, no fixed effects
m3_1 <- feols(as.formula(paste0("neg_spi_at ~ misalign_abs + ", ctrl)),
              data = dt, cluster = ~gvkey)

## (2) Firm FE only -- absorbs time-invariant firm characteristics
m3_2 <- feols(as.formula(paste0("neg_spi_at ~ misalign_abs + ", ctrl, " | gvkey")),
              data = dt, cluster = ~gvkey)

## (3) Firm + Year FE -- the main specification
m3_3 <- feols(as.formula(paste0("neg_spi_at ~ misalign_abs + ", ctrl, " | gvkey + fyear")),
              data = dt, cluster = ~gvkey)

## (4) Firm + Year FE + employee ideology, CEO turnover, and donor count controls
m3_4 <- feols(as.formula(paste0(
  "neg_spi_at ~ misalign_abs + emp_mean_cfscore + emp_sd_cfscore + ",
  "ceo_turnover + first_year_ceo + log_n_donors + ", ctrl, " | gvkey + fyear")),
  data = dt, cluster = ~gvkey)

## ── Build the table ──
vars_t3 <- c("misalign_abs", "emp_mean_cfscore", "emp_sd_cfscore",
             "ceo_turnover", "first_year_ceo", "log_n_donors",
             "lag_size", "lag_roa", "lag_leverage", "lag_tobinq",
             "lag_sale_growth", "lag_loss", "lag_capx_at", "lag_rd_at",
             "lag_aqc_indicator")
labels_t3 <- c("|Misalignment|", "Employee Mean CFscore", "Employee CFscore SD",
               "CEO Turnover", "First-Year CEO", "Log(No. Donors)",
               "Firm Size", "ROA", "Leverage", "Tobin's Q",
               "Sales Growth", "Loss", "Capex/Assets", "R&D/Assets", "Acquisition")

mods_t3 <- list(m3_1, m3_2, m3_3, m3_4)

rows_t3 <- do.call(rbind, lapply(seq_along(vars_t3), function(i) {
  make_row(vars_t3[i], labels_t3[i], mods_t3)
}))

fe_row <- c("Fixed Effects", "None", "Firm", "Firm + Year", "Firm + Year")
cl_row <- c("Cluster", rep("Firm", 4))
n_row <- c("Observations", sapply(mods_t3, function(m) fmt0(m$nobs)))
r2_row <- c("R-squared", sapply(mods_t3, function(m) fmt(get_r2(m), 3)))

tab3 <- as.data.frame(rbind(rows_t3, fe_row, cl_row, n_row, r2_row),
                       stringsAsFactors = FALSE)
names(tab3) <- c("", "(1)", "(2)", "(3)", "(4)")

kbl(tab3, booktabs = TRUE, align = c("l", rep("c", 4)),
    caption = "Baseline: Misalignment and Restructuring Charges",
    row.names = FALSE) %>%
  kable_styling(latex_options = c("hold_position"), font_size = 8,
                full_width = FALSE, position = "center") %>%
  row_spec(nrow(rows_t3), hline_after = TRUE) %>%
  footnote(
    general = "Dependent variable: Restructuring charges (= -SPI/AT; positive values indicate larger charges). Column (1) is pooled OLS. Column (2) adds firm fixed effects. Column (3) adds year fixed effects. Column (4) adds employee ideology, CEO turnover, and donor count controls. Standard errors clustered by firm in all columns. ***, **, * denote significance at the 1, 5, and 10 percent levels.",
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

\newpage

## Table 4: Recession Moderation of Misalignment Effect

The main result: tests whether the misalignment effect differs during
NBER recessions. The interaction `|Misalignment| x Recession` captures
the additional recession-period effect. Recession is year-level and
absorbed by year FE, so no standalone recession main effect is needed.

```{r table4}
## (1) OLS, no FE
m4_1 <- feols(as.formula(paste0(
  "neg_spi_at ~ misalign_abs + misalign_abs:recession + ", ctrl)),
  data = dt, cluster = ~gvkey)

## (2) Firm FE only
m4_2 <- feols(as.formula(paste0(
  "neg_spi_at ~ misalign_abs + misalign_abs:recession + ", ctrl, " | gvkey")),
  data = dt, cluster = ~gvkey)

## (3) Firm + Year FE -- main specification
m4_3 <- feols(as.formula(paste0(
  "neg_spi_at ~ misalign_abs + misalign_abs:recession + ", ctrl, " | gvkey + fyear")),
  data = dt, cluster = ~gvkey)

## (4) Firm + Year FE + employee ideology controls
m4_4 <- feols(as.formula(paste0(
  "neg_spi_at ~ misalign_abs + misalign_abs:recession + ",
  "emp_mean_cfscore + emp_sd_cfscore + ", ctrl, " | gvkey + fyear")),
  data = dt, cluster = ~gvkey)

## (5) Firm + Year FE + CEO turnover controls
m4_5 <- feols(as.formula(paste0(
  "neg_spi_at ~ misalign_abs + misalign_abs:recession + ",
  "ceo_turnover + first_year_ceo + ", ctrl, " | gvkey + fyear")),
  data = dt, cluster = ~gvkey)

mods_t4 <- list(m4_1, m4_2, m4_3, m4_4, m4_5)

vars_t4 <- c("misalign_abs", "misalign_abs:recession",
             "emp_mean_cfscore", "emp_sd_cfscore", "ceo_turnover", "first_year_ceo",
             "lag_size", "lag_roa", "lag_leverage", "lag_tobinq",
             "lag_sale_growth", "lag_loss", "lag_capx_at", "lag_rd_at",
             "lag_aqc_indicator")
labels_t4 <- c("|Misalignment|",
               "|Misalignment| \u00d7 Recession",
               "Employee Mean CFscore", "Employee CFscore SD",
               "CEO Turnover", "First-Year CEO",
               "Firm Size", "ROA", "Leverage", "Tobin's Q",
               "Sales Growth", "Loss", "Capex/Assets", "R&D/Assets", "Acquisition")

rows_t4 <- do.call(rbind, lapply(seq_along(vars_t4), function(i) {
  make_row(vars_t4[i], labels_t4[i], mods_t4)
}))

fe_row4 <- c("Fixed Effects", "None", "Firm", "Firm + Year",
             "Firm + Year", "Firm + Year")
cl_row4 <- c("Cluster", rep("Firm", 5))
n_row4 <- c("Observations", sapply(mods_t4, function(m) fmt0(m$nobs)))
r2_row4 <- c("R-squared", sapply(mods_t4, function(m) fmt(get_r2(m), 3)))

tab4 <- as.data.frame(rbind(rows_t4, fe_row4, cl_row4, n_row4, r2_row4),
                       stringsAsFactors = FALSE)
names(tab4) <- c("", "(1)", "(2)", "(3)", "(4)", "(5)")

kbl(tab4, booktabs = TRUE, align = c("l", rep("c", 5)),
    caption = "Recession Moderation of Misalignment Effect on Restructuring",
    row.names = FALSE) %>%
  kable_styling(latex_options = c("hold_position","scale_down"), font_size = 8,
                full_width = FALSE, position = "center") %>%
  row_spec(nrow(rows_t4), hline_after = TRUE) %>%
  footnote(
    general = paste0(
      "Dependent variable: Restructuring charges (-SPI/AT). ",
      "Recession = 1 during NBER recession years (2001, 2008--2009, 2020). ",
      "Recession is year-level and fully absorbed by year FE, so no standalone recession main effect is included. ",
      "Standard errors clustered by firm. ***, **, * denote significance at the 1, 5, and 10 percent levels."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

### Net effects from Column (3)

The non-recession coefficient is the base `|Misalignment|` effect. The
recession coefficient is the sum of `|Misalignment|` and
`|Misalignment| x Recession`, with the standard error computed via the
delta method.

```{r table4-net}
## Net effects: Non-recession vs Recession (from Column 3)
int_var <- "misalign_abs:recession"
r_base <- get_cp(m4_3, "misalign_abs")
r_rec  <- net_eff(m4_3, "misalign_abs", int_var, scale = 1)

net_df <- data.frame(
  Period = c("Non-recession", "Recession"),
  Effect = c(paste0(fmt4(r_base[1]), stars(r_base[3])),
             paste0(fmt4(r_rec[1]), stars(r_rec[3]))),
  SE = c(fmt4(r_base[2]), fmt4(r_rec[2])),
  pval = c(fmt4(r_base[3]), fmt4(r_rec[3])),
  stringsAsFactors = FALSE
)

kbl(net_df, booktabs = TRUE, align = c("l","r","r","r"),
    caption = "Net Effect of Misalignment by Recession Status (from Column 3)",
    row.names = FALSE,
    col.names = c("Period", "Net Effect", "SE", "p-value")) %>%
  kable_styling(latex_options = c("hold_position"), font_size = 9,
                full_width = FALSE, position = "center") %>%
  footnote(
    general = paste0(
      "Non-recession effect = coef(|Misalignment|). ",
      "Recession effect = coef(|Misalignment|) + coef(|Misalignment| x Recession). ",
      "***, **, * denote significance at the 1, 5, and 10 percent levels."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

\newpage

## Table 5: Donor Threshold Sensitivity

Re-estimates the main recession specification from Table 4 Column (3) at
progressively stricter donor thresholds (20+, 30+, 50+, 100+).

```{r table5}
thresholds <- c(20, 30, 50, 100)
f_sens <- paste0("neg_spi_at ~ misalign_abs + misalign_abs:recession + ", ctrl,
                 " | gvkey + fyear")

thresh_rows <- lapply(thresholds, function(thr) {
  ## Subset to firms with at least `thr` employee donors
  dts <- reg_data[!sic2 %in% c(49, 60:69) & n_emp_donors >= thr]
  dts[, lag_aqc_indicator := lag_acquisition]
  dts[, recession := nber_recession]
  m <- feols(as.formula(f_sens), data = dts, cluster = ~gvkey)

  r_base <- get_cp(m, "misalign_abs")
  r_int  <- get_cp(m, "misalign_abs:recession")

  data.frame(
    Threshold = paste0(thr, "+"),
    N = fmt0(m$nobs),
    Firms = fmt0(uniqueN(dts[!is.na(neg_spi_at)]$gvkey)),
    BaseCoef = paste0(fmt4(r_base[1]), stars(r_base[3])),
    BaseSE = fmt4(r_base[2]),
    IntCoef = paste0(fmt4(r_int[1]), stars(r_int[3])),
    IntSE = fmt4(r_int[2]),
    stringsAsFactors = FALSE
  )
})
thresh_df <- do.call(rbind, thresh_rows)

kbl(thresh_df, booktabs = TRUE, align = c("l","r","r","r","r","r","r"),
    caption = "Donor Threshold Sensitivity",
    row.names = FALSE,
    col.names = c("Threshold","N","Firms","Base Coef","Base SE",
                  "Int. Coef","Int. SE")) %>%
  kable_styling(latex_options = c("hold_position"), font_size = 9,
                full_width = FALSE, position = "center") %>%
  footnote(
    general = paste0(
      "Each row re-estimates the specification from Table 4, Column (3) on the subsample ",
      "with at least the indicated number of employee donors. ",
      "Base Coef is the |Misalignment| coefficient; Int. Coef is the interaction with Recession. ",
      "All models include firm + year FE, full controls, and firm-clustered SEs. ",
      "***, **, * denote significance at the 1, 5, and 10 percent levels."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

\newpage

## Table 6: CEO Ideology Level Test

Tests whether the *gap* (misalignment) matters beyond the CEO's own
ideology level. Column (2) adds `CEO CFscore` as a direct control.

```{r table6}
## (1) Baseline recession interaction
m6_1 <- feols(as.formula(paste0(
  "neg_spi_at ~ misalign_abs + misalign_abs:recession + ", ctrl, " | gvkey + fyear")),
  data = dt, cluster = ~gvkey)

## (2) Add CEO ideology level as a control
m6_2 <- feols(as.formula(paste0(
  "neg_spi_at ~ misalign_abs + ceo_cfscore + misalign_abs:recession + ", ctrl,
  " | gvkey + fyear")),
  data = dt, cluster = ~gvkey)

vars_t6 <- c("misalign_abs", "ceo_cfscore", "misalign_abs:recession")
labels_t6 <- c("|Misalignment|", "CEO CFscore",
               "|Misalignment| \u00d7 Recession")

mods_t6 <- list(m6_1, m6_2)
rows_t6 <- do.call(rbind, lapply(seq_along(vars_t6), function(i) {
  make_row(vars_t6[i], labels_t6[i], mods_t6)
}))

ctrl_row6 <- c("Controls", rep("Yes", 2))
fe_row6 <- c("Firm + Year FE", rep("Yes", 2))
cl_row6 <- c("Cluster", rep("Firm", 2))
n_row6 <- c("Observations", sapply(mods_t6, function(m) fmt0(m$nobs)))
r2_row6 <- c("R-squared", sapply(mods_t6, function(m) fmt(get_r2(m), 3)))

tab6 <- as.data.frame(rbind(rows_t6, ctrl_row6, fe_row6, cl_row6, n_row6, r2_row6),
                       stringsAsFactors = FALSE)
names(tab6) <- c("", "(1)", "(2)")

ceo_p <- get_cp(m6_2, "ceo_cfscore")[3]

kbl(tab6, booktabs = TRUE, align = c("l", rep("c", 2)),
    caption = "CEO Ideology Level Test",
    row.names = FALSE) %>%
  kable_styling(latex_options = c("hold_position"), font_size = 9,
                full_width = FALSE, position = "center") %>%
  row_spec(nrow(rows_t6), hline_after = TRUE) %>%
  footnote(
    general = paste0(
      "Dependent variable: Restructuring charges (-SPI/AT). Column (2) adds CEO CFscore as a direct control. ",
      "The misalignment coefficient and recession interaction remain stable while CEO CFscore is ",
      ifelse(ceo_p < 0.10, "significant", "insignificant"),
      " (p = ", fmt(ceo_p, 3), "), indicating that the ideological gap---not the CEO's ideology level---drives the effect. ",
      "Recession is year-level and absorbed by year FE. ",
      "***, **, * denote significance at the 1, 5, and 10 percent levels."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

\newpage

## Table 7: Asymmetry Test

Decomposes misalignment into two components: CEO more conservative than
employees (`ceo_right`) vs. CEO more liberal (`ceo_left`). Includes a
Wald test for whether the recession interactions differ.

```{r table7}
## Estimate asymmetric model with separate recession interactions
m7 <- feols(as.formula(paste0(
  "neg_spi_at ~ ceo_right + ceo_left + ",
  "ceo_right:recession + ceo_left:recession + ", ctrl, " | gvkey + fyear")),
  data = dt, cluster = ~gvkey)

vars_t7 <- c("ceo_right", "ceo_left",
             "ceo_right:recession", "ceo_left:recession",
             "lag_size", "lag_roa", "lag_leverage", "lag_tobinq",
             "lag_sale_growth", "lag_loss", "lag_capx_at", "lag_rd_at",
             "lag_aqc_indicator")
labels_t7 <- c("CEO More Conservative", "CEO More Liberal",
               "CEO More Cons. \u00d7 Recession",
               "CEO More Lib. \u00d7 Recession",
               "Firm Size", "ROA", "Leverage", "Tobin's Q",
               "Sales Growth", "Loss", "Capex/Assets", "R&D/Assets", "Acquisition")

rows_t7 <- do.call(rbind, lapply(seq_along(vars_t7), function(i) {
  make_row(vars_t7[i], labels_t7[i], list(m7))
}))

fe_row7 <- c("Firm + Year FE", "Yes")
cl_row7 <- c("Cluster", "Firm")
n_row7 <- c("Observations", fmt0(m7$nobs))
r2_row7 <- c("R-squared", fmt(get_r2(m7), 3))

tab7 <- as.data.frame(rbind(rows_t7, fe_row7, cl_row7, n_row7, r2_row7),
                       stringsAsFactors = FALSE)
names(tab7) <- c("", "(1)")

## Wald test: H0: coef(ceo_right:recession) == coef(ceo_left:recession)
bv <- coef(m7); Vv <- vcov(m7)
i_r <- match("ceo_right:recession", names(bv))
i_l <- match("ceo_left:recession", names(bv))
diff_b <- bv[i_r] - bv[i_l]
diff_se <- sqrt(Vv[i_r, i_r] + Vv[i_l, i_l] - 2 * Vv[i_r, i_l])
diff_p <- 2 * pt(abs(diff_b / diff_se), df = degrees_freedom(m7, "t"), lower.tail = FALSE)

kbl(tab7, booktabs = TRUE, align = c("l", "c"),
    caption = "Asymmetry Test: Conservative vs. Liberal CEO Misalignment",
    row.names = FALSE) %>%
  kable_styling(latex_options = c("hold_position"), font_size = 9,
                full_width = FALSE, position = "center") %>%
  row_spec(nrow(rows_t7), hline_after = TRUE) %>%
  footnote(
    general = paste0(
      "Dependent variable: Restructuring charges (-SPI/AT). ",
      "CEO More Conservative = max(signed misalignment, 0); ",
      "CEO More Liberal = max(-signed misalignment, 0). ",
      "Recession is year-level and absorbed by year FE. ",
      "Wald test H0: Cons. interaction = Lib. interaction: ",
      "Diff = ", fmt4(diff_b), ", SE = ", fmt4(diff_se), ", p = ", fmt(diff_p, 3), ". ",
      "***, **, * denote significance at the 1, 5, and 10 percent levels."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

\newpage

## Table 8: Leave-One-Recession-Out Robustness

Tests whether the recession interaction result is driven by any single
recession by sequentially excluding each recession period and
re-estimating.

```{r table8}
f_base <- paste0("neg_spi_at ~ misalign_abs + misalign_abs:recession + ", ctrl,
                 " | gvkey + fyear")

exclusions <- list(
  list(label = "Full sample",       excl = integer(0)),
  list(label = "Excl. 2001",        excl = 2001L),
  list(label = "Excl. 2008--2009",  excl = c(2008L, 2009L)),
  list(label = "Excl. 2020",        excl = 2020L),
  list(label = "Excl. 2020--2021",  excl = c(2020L, 2021L))
)

rob_rows <- lapply(exclusions, function(ex) {
  dts <- if (length(ex$excl) == 0) dt else dt[!fyear %in% ex$excl]
  m <- feols(as.formula(f_base), data = dts, cluster = ~gvkey)

  r_base <- get_cp(m, "misalign_abs")
  r_int  <- get_cp(m, "misalign_abs:recession")
  r_net  <- net_eff(m, "misalign_abs", "misalign_abs:recession", scale = 1)

  data.frame(
    Sample = ex$label,
    BaseCoef = paste0(fmt4(r_base[1]), stars(r_base[3])),
    BaseSE = fmt4(r_base[2]),
    IntCoef = paste0(fmt4(r_int[1]), stars(r_int[3])),
    IntSE = fmt4(r_int[2]),
    NetRec = paste0(fmt4(r_net[1]), stars(r_net[3])),
    NetSE = fmt4(r_net[2]),
    N = fmt0(m$nobs),
    stringsAsFactors = FALSE
  )
})
rob_df <- do.call(rbind, rob_rows)

kbl(rob_df, booktabs = TRUE, align = c("l","r","r","r","r","r","r","r"),
    caption = "Leave-One-Recession-Out Robustness",
    row.names = FALSE,
    col.names = c("Sample", "|Misalign|", "SE",
                  "Interaction", "SE",
                  "Net Recess.", "SE", "N")) %>%
  kable_styling(latex_options = c("hold_position"), font_size = 9,
                full_width = FALSE, position = "center") %>%
  footnote(
    general = paste0(
      "Dependent variable: Restructuring charges (-SPI/AT). ",
      "All models: |Misalignment| + |Misalignment| x Recession + controls | firm + year FE, clustered by firm. ",
      "|Misalign| is the non-recession effect. Interaction is the additional recession effect. ",
      "Net Recess. = |Misalign| + Interaction (total effect during recessions). ",
      "Each row excludes a different recession period. ",
      "***, **, * denote significance at the 1, 5, and 10 percent levels."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```
